---
title: Fuel Prices in Auckland
subtitle: STATS 220 Semester One, 2026
output: html_fragment
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)
```

```{css}

body {
  background: #f5f7f6;
  font-family: "Helvetica Neue", sans-serif;
  color: #2f3e3a;
  padding: 20px;
}

h1 {
  color: #ffffff;
  background-color: #2f6f62; /* foam sea green */
  padding: 18px;
  text-align: center;
  font-weight: 600;
  border-radius: 6px;
  letter-spacing: 1px;
}

h2 {
  color: #ffffff;
  background-color: #4f7f6f; /* moss green */
  padding: 12px;
  font-weight: 600;
  border-radius: 4px;
  margin-top: 30px;
}

h3 {
  color: #2f6f62;
  font-weight: 600;
}

p {
  font-size: 16px;
  line-height: 1.6;
  color: #394744;
}

li {
  font-size: 16px;
}

.section {
  background-color: #ffffff;
  padding: 18px;
  border-radius: 6px;
  margin-bottom: 25px;
  border-left: 5px solid #4f7f6f;
  box-shadow: 0px 2px 6px rgba(0,0,0,0.05);
}

img {
  display: block;
  margin-left: auto;
  margin-right: auto;
  margin-top: 20px;
  margin-bottom: 20px;
  border-radius: 4px;
}

table {
  border-collapse: collapse;
  width: 100%;
  margin-top: 20px;
}

th {
  background-color: #2f6f62;
  color: white;
  padding: 10px;
  text-align: left;
}

td {
  padding: 10px;
  border-bottom: 1px solid #ddd;
}

tr:hover {
  background-color: #f1f4f3;
}

```

# Location Summary

The dataset shows fuel prices collected across different Auckland locations.

```{r}
library("tidyverse")

csv_file <- "https://docs.google.com/spreadsheets/d/e/2PACX-1vTElqKsgASNluXUaano0vLYorvYaBtoDSF2gCqncpWcQAzg2BnE7kft1m-gm_eIMdFa0fGTrocxx_Ou/pub?output=csv"

logged_data <- read_csv(csv_file)

logged_data <- logged_data %>%
  rename(
    `Date` = `Enter the date`,
    `Time` = `Enter the time`,
    `Fuel Brand` = `Select the fuel station brand`,
    `Location` = `Where is the fuel station located? Only enter the suburb/town/district.`
  )

# Location and Time data #

logged_data %>%
  ggplot(aes(x = `Location`)) +
  geom_bar() +
  theme_minimal() +
  labs(
    title = "Frequency of passing Locations",
    x = "Locations",
    y = "Count"
  )




# Unleaded 91 data

unleaded_91 <- logged_data$`Unleaded 91`
average_91 <- sum(unleaded_91) / length(unleaded_91)
average_91 <- round(average_91, 2)

# Summary values for Unleaded 91
min_91 <- min(logged_data$`Unleaded 91`, na.rm = TRUE)
max_91 <- max(logged_data$`Unleaded 91`, na.rm = TRUE)
avg_91 <- round(mean(logged_data$`Unleaded 91`, na.rm = TRUE), 2)

```

# Summary of Unleaded 91 Prices

The distribution of Unleaded 91 prices shows that the **minimum price** is $`r min_91`, the **maximum price** is $`r max_91`, and the **average price** is $`r avg_91`. This indicates that prices are fairly consistent across locations.


```{r}
# bar graph of Unleaded 91 prices
logged_data %>%
  ggplot(aes(x = factor(`Unleaded 91`))) +
  geom_bar(fill = "darkolivegreen") +
  theme_minimal() +
  labs(
    title = "Unleaded 91 Price Distribution",
    x = "Unleaded 91 Price",
    y = "Count"
  )


# Premium 95 data

premium_95 <- logged_data$`Premium 95`
filtered_95 <- premium_95[!is.na(premium_95)]
average_95 <- sum(premium_95, na.rm = TRUE) / length(filtered_95)
average_95 <- round(average_95, 2)

# Summary values for Premium 95
min_95 <- min(logged_data$`Premium 95`, na.rm = TRUE)
max_95 <- max(logged_data$`Premium 95`, na.rm = TRUE)
avg_95 <- round(mean(logged_data$`Premium 95`, na.rm = TRUE), 2)

```


# Summary of Premium 95 Prices

The distribution of Premium 95 prices shows that the **minimum price** is $`r min_95`, the **maximum price** is $`r max_95`, and the **average price** is $`r avg_95`. This suggests that Premium 95 prices are slightly higher than Unleaded 91 but remain fairly consistent across locations.

```{r}

# Summary of Premium 95 Prices

logged_data %>%
  ggplot(aes(x = `Premium 95`)) +
  geom_bar(fill = "palegreen3") +
  theme_minimal() +
  labs(
    title = "Premium 95 Price Distribution",
    x = "Premium 95 Price",
    y = "Count"
  )


# Premium 98 data

premium_98 <- logged_data$`Premium 98`
filtered_98 <- premium_98[!is.na(premium_98)]
average_98 <- sum(premium_98, na.rm = TRUE) / length(filtered_98)
average_98 <- round(average_98, 2)

# Summary values for Premium 98
min_98 <- min(logged_data$`Premium 98`, na.rm = TRUE)
max_98 <- max(logged_data$`Premium 98`, na.rm = TRUE)
avg_98 <- round(mean(logged_data$`Premium 98`, na.rm = TRUE), 2)

```

# Summary of Premium 98 Prices

The distribution of Premium 98 prices shows that the **minimum price** is $`r min_98`, the **maximum price** is $`r max_98`, and the **average price** is $`r avg_98`. This suggests that Premium 98 prices are generally higher than other petrol kinds, but remain fairly similar across locations.

```{r}

logged_data %>%
  ggplot(aes(x = `Premium 98`)) +
  geom_bar(fill = "seagreen") +
  theme_minimal() +
  labs(
    title = "Premium 98 Price Distribution",
    x = "Premium 98 Price",
    y = "Count"
  )

  
# Overall data

# Pivot data long for easy calculation
fuel_long <- logged_data %>%
  select(Location, `Unleaded 91`, `Premium 95`, `Premium 98`) %>%
  pivot_longer(
    cols = c(`Unleaded 91`, `Premium 95`, `Premium 98`),
    names_to = "Fuel Type",
    values_to = "Price"
  )

# Get min and max location for each fuel type
fuel_summary <- fuel_long %>%
  group_by(`Fuel Type`) %>%
  summarise(
    MinPrice = min(Price, na.rm = TRUE),
    MaxPrice = max(Price, na.rm = TRUE),
    MinLoc = Location[which.min(Price)],
    MaxLoc = Location[which.max(Price)]
  )

```


# Summary of Prices by Location

```{r}

logged_data %>%
  select(Location, `Unleaded 91`, `Premium 95`, `Premium 98`) %>%
  pivot_longer(
    cols = c(`Unleaded 91`, `Premium 95`, `Premium 98`),
    names_to = "Fuel Type",
    values_to = "Price"
  ) %>%
  # filter and create graph from data
  group_by(Location, `Fuel Type`) %>%
  summarise(avg_price = mean(Price, na.rm = TRUE)) %>%
  ggplot(aes(x = Location, y = avg_price, fill = `Fuel Type`)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  # calculate average prices
  geom_text(aes(label = round(avg_price, 2)),
            position = position_dodge(width = 0.8),
            vjust = -0.3,
            size = 3) +
  # choose colours
  scale_fill_manual(values = c(
    "Unleaded 91" = "#4f7f6f",
    "Premium 95" = "#2f6f62",
    "Premium 98" = "#8faea3"
  )) +
  # add labels
  labs(
    title = "Average Petrol Prices by Location",
    x = "Location",
    y = "Average Price ($)",
    fill = "Fuel Type"
  ) +
  # make labels easier to read
  coord_cartesian(ylim = c(2.9, 4.1)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

